---
title: "Working with Dates and regex"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(dplyr)
library(ggplot2)
```

# Purpose
Explore BIKETOWN data looking at mean miles travelled per week during the summer of 2018. 06/18-09/18

```{r pull_data}
source("/users/kalemattias/Desktop/plot/analyze_biketown.R")
get_data(start = "06/2018", end = "09/2018")

setwd("/users/kalemattias/Desktop/plot/data/biketown/")

folder <- "/users/kalemattias/Desktop/plot/data/biketown/"
filenames <- list.files(path = folder, pattern = "*.csv", all.files = FALSE,
                        full.names = FALSE, recursive = FALSE, 
                        ignore.case = FALSE)
read_csv_filename <- function(filenames) {
  ret <- read.csv(filenames, stringsAsFactors = FALSE, 
                  strip.white = TRUE, na.strings = "")
ret$Source <- filenames
ret
}
bike_raw <- plyr::ldply(filenames, read_csv_filename)

```
```{r data wrangling}
str(bike_raw)

bike_dfl <- bike_raw %>% 
  mutate(start.datetime = paste(StartDate, StartTime, sep = " "),
         end.datetime = paste(EndDate, EndTime, sep = " "))
str(bike_dfl)

# convert start and end date time into date-time format
# make sure to include time zone

# check acceptable timezone values
OlsonNames()

bike_dfl$start.datetime <- mdy_hm(bike_dfl$start.datetime, tz= "America/Los_Angeles")
bike_dfl$end.datetime <- mdy_hm(bike_dfl$end.datetime, tz= "America/Los_Angeles")
str(bike_dfl)

#convert duration into useable format
bike_dfl$Duration <- hms(bike_dfl$Duration)
#error thrown after converting to duration indicating NAs, 
#check raw data frame for NAs in duration column

sum(is.na(bike_raw$Duration))
```

"lubridate" has three synonomous functions but they define very different actions:
1. duration(): time spans are in united of seconds, there is no start date.
2. interval(): also measrued in unites of seconds, and provides the diff between start and end.
3. period(): measures time span in units larger than seconds, this is handy when acounting for DST and leap years.

``` {r duration_calc}
#calc interval
bike_dfl$interval <- interval(bike_dfl$start.datetime, bike_dfl$end.datetime)
head(bike_dfl$interval)

# calc duration
bike_dfl$duration.all <- as.duration(bike_dfl$interval)
head(bike_dfl$duration.all)

#calc period
bike_dfl$period <- as.period(bike_dfl$duration.all)
head(bike_dfl$period)

# floor_date() is hlepful for aggregating
bike_wkagg <- bike_dfl %>% 
  mutate(week.datetime = floor_date(start.datetime, unit = "week"))
head(bike_wkagg$week.datetime)

# calc mean distance travelled in miles per week
#gdp_continent <- gapminder %>% 
  #group_by(continent) %>% 
  #summarise(mean_gdp = mean(gdpPercap), mean_lifeExp = mean(lifeExp))

wk_avg_dist <- bike_wkagg %>%
  group_by(week.datetime) %>% 
  summarise(weekly.mean_dist = mean(Distance_Miles))
View(wk_avg_dist)

wk_avg_dist_fig <- wk_avg_dist %>% 
  ggplot(aes(x = as.Date(week.datetime), y = weekly.mean_dist)) +
  geom_bar(stat = "identity", fill = "orange") +
  scale_x_date(date_breaks = "1 week") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
wk_avg_dist_fig
```
